Formal Verification Framework Based on Symbolic Execution for Smart Contract

### 2023-12-12 papers/static-checker/ecoop2024/ecoop2024.tex
Write DSL section

In this work, we present the design of a domain-specific language for specifying smart contract properties and demonstrate the integration of symbolic execution with an SMT solver for formal verification of these properties.
\section{Symbolic Execution Model}
\label{sec:symbolic-execution-model}
\subsection{System Model}
Michelson serves as the low-level smart contract language for the Tezos blockchain platform. The processing and manipulation of data during the execution of smart contracts are managed using a stack. The representation of a stack can be described as follows:
\begin{itemize}
\item[]  \EMPTYSTACK\ represents the empty stack.
\item[]  \HEAD\ \STACKCONCAT\ \NTAIL\ denotes a stack with the top element as \HEAD\ and the remaining stack as \NTAIL.
\end{itemize}
\begin{figure} []
\begin{align*}
T, U &::= \\
   &\Mid\ \langle \text{comparable type}\rangle \\
   &\Mid\ \text{option}\ \langle\text{type}\rangle \\
   &\Mid\ \text{list}\ \langle\text{type}\rangle \\
   &\Mid\ \text{set}\ \langle\text{comparable type}\rangle \\
   &\Mid\ \text{operation} \\
   &\Mid\ \text{contract}\ \langle\text{type}\rangle \\
   &\Mid\ \text{ticket}\ \langle\text{comparable type}\rangle \\
   &\Mid\ \text{pair}\ \langle\text{type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{or}\ \langle\text{type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{lambda}\ \langle\text{type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{map}\ \langle\text{comparable type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{big-map}\ \langle\text{comparable type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{bls12-381-g1} \\
   &\Mid\ \text{bls12-381-g2} \\
   &\Mid\ \text{bls12-381-fr} \\
   &\Mid\ \text{sapling-transaction}\ \langle\text{natural number constant}\rangle \\
   &\Mid\ \text{sapling-state}\ \langle\text{natural number constant}\rangle \\
   &\Mid\ \text{chest} \\
   &\Mid\ \text{chest-key} \\
\langle\text{comparable type}\rangle ::= \\
   &\Mid\ \text{unit} \\
   &\Mid\ \text{never} \\
   &\Mid\ \text{bool} \\
   &\Mid\ \text{int}\\
   &\Mid\ \text{nat}\\
   &\Mid\ \text{string}\\
   &\Mid\ \text{chain-id}\\
   &\Mid\ \text{bytes}\\
   &\Mid\ \text{mutez}\\
   &\Mid\ \text{key-hash}\\
   &\Mid\ \text{key}\\
   &\Mid\ \text{signature}\\
   &\Mid\ \text{timestamp}\\
   &\Mid\ \text{address}\\
   &\Mid\ \text{tx-rollup-l2-address}\\
   &\Mid\ \text{option}\ \langle\text{comparable type}\rangle\\
   &\Mid\ \text{or}\ \langle\text{comparable type}\rangle \langle\text{comparable type}\rangle\\
   &\Mid\ \text{pair}\ \langle\text{comparable type}\rangle \langle\text{comparable type}\rangle \DOT \\
\end{align*}
\caption{Types}
\label{fig:type}
\end{figure}
Each piece of data in Michelson is assigned a specific type, and the language restricts programmers from introducing custom types. The complete list of types can be found in Figure \ref{fig:type}. To model the system, we use a stack in which each element is a pair consisting of term and its corresponding type. Terms are defined in Figure \ref{fig:term}. The stack is denoted as \STACK\ = (\TermOne, \TYF) \STACKCONCAT\ (\TermTwo, \TYS) \STACKCONCAT\ \DOT\ \STACKCONCAT\ \EMPTYSTACK\ where elements are types paired with terms.

@@

The smart contract program takes a stack as input, consisting of a single pair where the first element is an input value referred to as parameter denoted as \VPAR and the second element represents the content of the storage space denoted as \VSTORAGE. The program produces a stack as output, which is a single pair. The first element of this pair is the list of internal operations that the program intends to emit (\VOPERATIONLIST), and the second element is the new contents of the storage space.  Let's define the initial stack as  \SINIT\   = (\KPAIR\ \VPAR\ \VSTORAGE, \TPAIR\ \TYF\ \TYS) \STACKCONCAT\ \EMPTYSTACK, where \VPAR\ and \VSTORAGE\ denote terms representing the parameter and the storage, and \TYF\ and \TYS\ are their corresponding types. Similarly, designate the final stack as \SFINAL\   = (\PAIR\ \VOPERATIONLIST\ \VSTORAGE, \TPAIR\ (\TOPERATIONLIST) \TYS) \STACKCONCAT\ \EMPTYSTACK, where \VOPERATIONLIST\ repesents a operation list.
\begin{figure}[]
\begin{align*}
\text{t} &::= \\
   &\Mid\ \langle \text{variable} \rangle \\
   &\Mid\ \langle \text{account constant} \rangle \\
   &\Mid\ \langle \text{int constant} \rangle \\
   &\Mid\ \langle \text{string constant} \rangle \\
   &\Mid\ \langle \text{byte sequence constant} \rangle \\
   &\Mid\ \UNIT \\
   &\Mid\ \NEVER \\
   &\Mid\ \TRUE \\
   &\Mid\ \FALSE \\
   &\Mid\ \PAIR\ \text{t1 t2}\\
   &\Mid\ \LEFT\ \text{t}\\
   &\Mid\ \RIGHT\ \text{t}\\ 
   &\Mid\ \SOME\ \text{t}\\
   &\Mid\ \NONE \\
   &\Mid\ \text{\{t ; ... \}}\\
   &\Mid\ \text{\{ Elt t1 t2 ; ... \}}\\
   &\Mid\ \{ \langle\text{instruction}\rangle; ... \}   \\
\langle \text{variable} \rangle &::= \\ 
   &\Mid\ \VAR\\
\langle \text{account constant} \rangle &::= \\ 
   &\Mid\ \text{balance} \\
   &\Mid\ \text{amount} \\
   &\Mid\ \text{sender} \\
   &\Mid\ \text{source} \\
   &\Mid\ \text{now} \\
   &\Mid\ \text{level} \\
   &\Mid\ \text{chain-id} \\
   &\Mid\ \text{self}  \\
   &\Mid\ \text{self-address}  \\
   &\Mid\ \text{total-voting-power}  \\
   &\Mid\ \text{voting-power}  \\
\langle\text{instruction}\rangle &::= \\
  &\Mid\ \DROP \\
  &\Mid\ \DROP \langle\text{natural number constant}\rangle \\
  &\text{...}
\end{align*}
\caption{Terms}
\label{fig:term}
\end{figure}
\subsection{Rules}
Each instruction is defined by a rule on the system model. The rule semantic is defined by several kinds of transitions:
\begin{enumerate}
\item \ExprTrans\ single-step evaluation of an expression in a system state,
\item \StateTrans\ internal transitions of a state, which is non-deterministic  as shown for loop instructions, such as \LOOP\ and \LOOP-LEFT,
\item \SystemTrans\ symbolic system transitions.
\end{enumerate}

@@

\begin{mathpar}
\inferrule[]
  { \NEG\ \PREDICATE
  }{
  \{[\INSTRUCTION, \STACK, \PREDICATE]\} \cup \SYSTEM \SystemTrans \SYSTEM}
\end{mathpar}
Michelson instructions are categorized based on their functions. These include such as control structures, stack manipulation, arithmetic operations, boolean operations, cryptographic operations, and operations on data structures. In our model, we classified them as one-step, multi-step, blockchain, cryptographic, branch, and loop instructions. This categorization reflects how these instructions are handled within our specific model.
\subsubsection{One-step Instructions}
One-step instructions have a localized effect, modifying the system state directly without branching a new state, and they end the execution after a single operation. The ABS rule exemplifies such an instruction, operating on the first element of the stack. This element must be of type int, and the instruction returns the positive value of its current value, treated as a natural number.
%ABS
\begin{mathpar}
\inferrule[ABS]
  {
  }
  {
    [(\ABS; \INSTRUCTION), (\StackOne, \TINT) \STACKCONCAT \STACK,
    \PREDICATE]
    \StateTrans 
    [\INSTRUCTION, (\VariableX, \TNAT) \STACKCONCAT \STACK,
    \PREDICATE \wedge (\StackOne \ge 0 \Rightarrow \VariableX =
    \StackOne) \wedge (\StackOne <0 \Rightarrow \VariableX = -\StackOne)]
 }
\end{mathpar}
Another example is the COMPARE instruction for natural numbers, which compares the first two elements of the stack, both having a natural number type. It returns a result of either -1, 0, or 1 based on the comparison of their values.
% COMPARE
\begin{mathpar}
\inferrule[COMPARE]
  {
  }
  {
    [(\COMPARE ; \INSTRUCTION), (\StackOne, \TNAT) \STACKCONCAT (\StackTwo, \TNAT)
    \STACKCONCAT \STACK, \PREDICATE ]
    \SystemTrans \\
    [\INSTRUCTION, (\VariableX, \TINT) \STACKCONCAT \STACK, \PREDICATE
    \wedge\ (\StackOne > \StackTwo \Leftrightarrow \VariableX = 1)
    \wedge\ (\StackOne = \StackTwo \Leftrightarrow \VariableX = 0) 
    \wedge\ (\StackOne < \StackTwo \Leftrightarrow \VariableX = -1)]
    }
\end{mathpar}
\subsubsection{Multi-step Instructions}
Multi-step instructions involve the execution of a sub-sequence of instructions before returning to the main execution. The rule below models the \EXEC\ instruction, which executes a function specified as a sequence of instructions denoted by \INSTRUCTIONONE\ in the rule. This instruction applies the code of the function to the first element of the stack and then places the result back into the main execution.
\begin{mathpar}
  \inferrule[\EXEC]
  {
    [\INSTRUCTIONONE, (\StackOne, \TYF) \STACKCONCAT \EMPTYSTACK, 
    Q]
    \StateTrans^*
    [ \EMPTYSTACK, (\StackOne', \TYS) \STACKCONCAT Q']
  }{
     [(\EXEC; \INSTRUCTION),   (\{\INSTRUCTIONONE\}, \TYF\ \rightarrow\ \TYS) \STACKCONCAT
    (\StackOne, \TYF) \STACKCONCAT \STACK, \PREDICATE \wedge Q] 
    \StateTrans \
    [ \INSTRUCTION, (\StackOne', \TYS) \STACKCONCAT \STACK,
    \PREDICATE \wedge Q']}
\end{mathpar}

The \DIP\ \N\ instruction delves into the first \N\ elements of a stack, applies the sequence of instructions denoted as \INSTRUCTIONONE\ to the remaining part of the stack, modifies a part of the predicate and subsequently pushes the result back into the main execution.
%DIP n
\begin{mathpar}
\inferrule[\DIP\ \N]
  { 
     \FLEN(\A) \EQUAL\ \N \\ [\INSTRUCTIONONE,  \B, Q]
    \StateTrans^*
    [\EMPTYSTACK,  \B_1, Q']
  }
  {[(\DIP\ \N\ \INSTRUCTIONONE; \INSTRUCTION), \A\ \At\ \B, \PREDICATE \wedge Q] \StateTrans 
[(\INSTRUCTION), \A\ \At\ \B_1, \PREDICATE \wedge Q']}
\end{mathpar}
\subsubsection{Blockchain Instructions}
 Blockchain instructions make implicit use of data from an execution environment, derived from the current state of the blockchain (the context), from the transaction that triggered the contract, and from the block containing this transaction. We model these data as symbolic values, which satisfis some constrants. For example, the AMOUNT instruction adds to the stack the amount of tokens that sent together with the transcation which calls this smart contract. Here, the amount is modeled as a symbolic value donated as \VAMOUNT, which has \TMUTEZ\ type, which has a nature number value.

\begin{mathpar}
\inferrule[AMOUNT]
  {
  }
  {[(\AMOUNT; \INSTRUCTION), \STACK, \PREDICATE] \StateTrans \
[\INSTRUCTION, (\VAMOUNT, \TMUTEZ) \STACKCONCAT\STACK, \PREDICATE}
\end{mathpar}

\subsubsection{Cryptographic Instructions}
Cryptographic instructions in Michelson utilize built-in functions that manage cryptographic features. These functions are symbolically modeled in our framework, specifying their inputs and outputs. For instance, the instruction \SHA\ computes the cryptographic hash of the top of the stack using the SHA-256 cryptographic hash function. The function SHA-256 is symbolically represented as the function  \FSHA\  that takes an input of type \TBYTE\ and returns another byte value, constrained to a length of 256.
%HASH-KEY
\begin{mathpar}
\inferrule[\SHA]
  {
  }
  {[(\SHA; \INSTRUCTION), (\StackOne, \TBYTE) \STACKCONCAT\STACK, \PREDICATE] \StateTrans \
[\INSTRUCTION, (\VariableX, \TBYTE) \STACKCONCAT\STACK, \PREDICATE \Wedge\ (\VariableX\ = \FSHA(\StackOne))]}
\end{mathpar}
\subsubsection{Branch Instructions}
The Michelson language incorporates conditional branching, enabling the execution of specific code segments based on certain criteria. One of the instructions facilitating this is the \IF\ \{\} \{\} construct. For example, the \IF\ \{ \INSTRUCTIONONE\ \} \{ \INSTRUCTIONTWO\  \} instruction allows the creation of branches in execution. It takes two sequences as arguments. Expecting a boolean value at the top of the stack, it consumes this top element. If the boolean is True, it executes the first provided sequence; otherwise, it executes the second sequence.
%IF
\begin{mathpar}
  \inferrule[IF-true]
  {  
  }{
    [(\IF\ \INSTRUCTIONONE\  \INSTRUCTIONTWO; \INSTRUCTION),
    (\StackOne, \TBOOL) \STACKCONCAT\STACK, \PREDICATE]
    \StateTrans\
    [\INSTRUCTIONONE, \STACK, \PREDICATE\ \Wedge\ \StackOne]
  }

@@

Another branch instruction is the IF-LEFT instruction, which is used for conditional execution based on the left or right branch of a tagged union (sum) type. 
%IF-LEFT-LEFT
\begin{mathpar}
  \inferrule[IF-LEFT-left]
  {  
  }{
    [(\IFLEFT\ \INSTRUCTIONONE\ \INSTRUCTIONTWO; \INSTRUCTION),
    (\StackOne, \TOR\ \TYF\ \TYS) \STACKCONCAT \STACK, \PREDICATE]
    \StateTrans \
    [\INSTRUCTIONONE, (\VariableX, \TYF) \STACKCONCAT\STACK,
    \PREDICATE \wedge (\StackOne\ \EQUAL\ \LEFT\ \VariableX)]
  }
\end{mathpar}

%IF-LEFT-RIGHT
\begin{mathpar}
  \inferrule[IF-LEFT-right]
  {  
  }{
    [(\IFLEFT\ \INSTRUCTIONONE\  \INSTRUCTIONTWO; \INSTRUCTION),
    (\StackOne, \TOR\ \TYF\ \TYS) \STACKCONCAT \STACK, \PREDICATE]
    \StateTrans \
    [\INSTRUCTIONTWO, (\VariableX, \TYS) \STACKCONCAT\STACK, \PREDICATE \wedge (\StackOne\ \EQUAL\ \RIGHT\ \VariableX))]
  }
\end{mathpar}
These above rules collectively define the behavior of the IF-LEFT instruction based on the left or right branch of a tagged union type. This rule applies when the top element of the stack is a tagged union of type \TOR\ \TYF\ \TYS (either of type \TYF\ or \TYS), and the IF-LEFT instruction is followed by instructions \INSTRUCTIONONE\ and \INSTRUCTIONTWO. The first rule apply with the condition that the top element of the stack is of the form Left \VariableX\ (indicating the left branch of the union), it transitions to execute the instructions \INSTRUCTIONONE\ with the stack modified accordingly and the predicate is updated to include the condition that the top element of the stack is equal to Left \VariableX.  Similerly, the second rule apply with the condition that  the top element is of the form Right \VariableX\ (indicating the right branch of the union), it transitions to execute the instructions \INSTRUCTIONTWO\ with the stack modified accordingly and the predicate is updated to include the condition that the top element of the stack is equal to Right \VariableX.

\subsubsection{Loop Instructions}
In Michelson, loops are fundamental constructs that repeatedly execute a set of instructions until a specific condition is met. There are various loop instructions, including ITER, MAP, CONCAT, SIZE, LOOP, and LOOP-LEFT, enable the manipulation and traversal of different data structures like lists, sets, maps, and the stack itself.

Michelson, similar to other programming languages, classifies loop instructions into two main types:
\begin{itemize}
\item \KFOR\ loops: These are sequence-based loops, exemplified by ITER, MAP, CONCAT and SIZE.
\item \KWHILE\ loops: These are predicates-based loops, illustrated by LOOP and LOOP-LEFT.
\end{itemize}
To illustrate the structures of  loops, we can draw a parallel with OCaml syntax:
\begin{verbatim}
while <loop_condition> do
   <loop_body>
done
\end{verbatim}
This \KWHILE\  loop structure involves a loop condition that is evaluated at the beginning of each iteration. If the condition holds true, the loop body is executed, and the process repeats until the condition becomes false.  In Michelson, this pattern is evident in instructions like LOOP and LOOP-LEFT, which rely on predicate-based conditions for their iteration.
\begin{verbatim}
LOOP {<loop-body>}
\end{verbatim}
Additionally, Michelson features a \KFOR\ loop, where the index runs from the initial value to the final value:
\begin{verbatim}
for index = <initial_value> to <final_value> do
  <loop_body>
done
\end{verbatim}
This is represented by the ITER instruction:
\begin{verbatim}
ITER {<loop-body>}
\end{verbatim}

>>>>

In this work, we present the design of a domain-specific language for specifying smart contract properties and demonstrate the integration of symbolic execution with an SMT solver for formal verification of these properties.
\section{Symbolic Execution Model}
\label{sec:symbolic-execution-model}
\subsection{System Model}
Michelson serves as the low-level smart contract language for the Tezos blockchain platform. The processing and manipulation of data during the execution of smart contracts are managed using a stack. The representation of a stack can be described as follows:
\begin{itemize}
\item[]  \ensuremath{[\ ]} represents the empty stack.
\item[]  \ensuremath{hd :: tl} denotes a stack with the top element as \ensuremath{hd} and the remaining stack as \ensuremath{tl}.
\end{itemize}
\begin{figure} []
\begin{align*}
T, U &::= \\
   &\Mid\ \langle \text{comparable type}\rangle \\
   &\Mid\ \text{option}\ \langle\text{type}\rangle \\
   &\Mid\ \text{list}\ \langle\text{type}\rangle \\
   &\Mid\ \text{set}\ \langle\text{comparable type}\rangle \\
   &\Mid\ \text{operation} \\
   &\Mid\ \text{contract}\ \langle\text{type}\rangle \\
   &\Mid\ \text{ticket}\ \langle\text{comparable type}\rangle \\
   &\Mid\ \text{pair}\ \langle\text{type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{or}\ \langle\text{type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{lambda}\ \langle\text{type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{map}\ \langle\text{comparable type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{big-map}\ \langle\text{comparable type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{bls12-381-g1} \\
   &\Mid\ \text{bls12-381-g2} \\
   &\Mid\ \text{bls12-381-fr} \\
   &\Mid\ \text{sapling-transaction}\ \langle\text{natural number constant}\rangle \\
   &\Mid\ \text{sapling-state}\ \langle\text{natural number constant}\rangle \\
   &\Mid\ \text{chest} \\
   &\Mid\ \text{chest-key} \\
\langle\text{comparable type}\rangle ::= \\
   &\Mid\ \text{unit} \\
   &\Mid\ \text{never} \\
   &\Mid\ \text{bool} \\
   &\Mid\ \text{int}\\
   &\Mid\ \text{nat}\\
   &\Mid\ \text{string}\\
   &\Mid\ \text{chain-id}\\
   &\Mid\ \text{bytes}\\
   &\Mid\ \text{mutez}\\
   &\Mid\ \text{key-hash}\\
   &\Mid\ \text{key}\\
   &\Mid\ \text{signature}\\
   &\Mid\ \text{timestamp}\\
   &\Mid\ \text{address}\\
   &\Mid\ \text{tx-rollup-l2-address}\\
   &\Mid\ \text{option}\ \langle\text{comparable type}\rangle\\
   &\Mid\ \text{or}\ \langle\text{comparable type}\rangle \langle\text{comparable type}\rangle\\
   &\Mid\ \text{pair}\ \langle\text{comparable type}\rangle \langle\text{comparable type}\rangle \DOT \\
\end{align*}
\caption{Types}
\label{fig:type}
\end{figure}
Each piece of data in Michelson is assigned a specific type, and the language restricts programmers from introducing custom types. The complete list of types can be found in Figure \ref{fig:type}. To model the system, we use a stack in which each element is a pair consisting of term and its corresponding type. Terms are defined in Figure \ref{fig:term}. The stack is denoted as \STACK\ = (\TermOne, \TYF) \STACKCONCAT\ (\TermTwo, \TYS) \STACKCONCAT\ \DOT\ \STACKCONCAT\ \EMPTYSTACK\ where elements are types paired with terms.

@@

The smart contract program takes a stack as input, consisting of a single pair where the first element is an input value referred to as parameter denoted as \VPAR and the second element represents the content of the storage space denoted as \VSTORAGE. The program produces a stack as output, which is a single pair. The first element of this pair is the list of internal operations that the program intends to emit (\VOPERATIONLIST), and the second element is the new contents of the storage space.  Let's define the initial stack as  \SINIT\   = (\KPAIR\ \VPAR\ \VSTORAGE, \TPAIR\ \TYF\ \TYS) \STACKCONCAT\ \EMPTYSTACK, where \VPAR\ and \VSTORAGE\ denote terms representing the parameter and the storage, and \TYF\ and \TYS\ are their corresponding types. Similarly, designate the final stack as \SFINAL\   = (\PAIR\ \VOPERATIONLIST\ \VSTORAGE, \TPAIR\ (\TOPERATIONLIST) \TYS) \STACKCONCAT\ \EMPTYSTACK, where \VOPERATIONLIST\ repesents a operation list.
\begin{figure}[]
\begin{align*}
\text{t} &::= \\
   &\Mid\ \langle \text{variable} \rangle \\
   &\Mid\ \langle \text{account constant} \rangle \\
   &\Mid\ \langle \text{int constant} \rangle \\
   &\Mid\ \langle \text{string constant} \rangle \\
   &\Mid\ \langle \text{byte sequence constant} \rangle \\
   &\Mid\ \UNIT \\
   &\Mid\ \NEVER \\
   &\Mid\ \TRUE \\
   &\Mid\ \FALSE \\
   &\Mid\ \PAIR\ \text{t1 t2}\\
   &\Mid\ \LEFT\ \text{t}\\
   &\Mid\ \RIGHT\ \text{t}\\ 
   &\Mid\ \SOME\ \text{t}\\
   &\Mid\ \NONE \\
   &\Mid\ \text{\{t ; ... \}}\\
   &\Mid\ \text{\{ Elt t1 t2 ; ... \}}\\
   &\Mid\ \{ \langle\text{instruction}\rangle; ... \}   \\
\langle \text{variable} \rangle &::= \\ 
   &\Mid\ \VAR\\
\langle \text{account constant} \rangle &::= \\ 
   &\Mid\ \text{balance} \\
   &\Mid\ \text{amount} \\
   &\Mid\ \text{sender} \\
   &\Mid\ \text{source} \\
   &\Mid\ \text{now} \\
   &\Mid\ \text{level} \\
   &\Mid\ \text{chain-id} \\
   &\Mid\ \text{self}  \\
   &\Mid\ \text{self-address}  \\
   &\Mid\ \text{total-voting-power}  \\
   &\Mid\ \text{voting-power}  \\
\langle\text{instruction}\rangle &::= \\
  &\Mid\ \DROP \\
  &\Mid\ \DROP \langle\text{natural number constant}\rangle \\
  &\text{...}
\end{align*}
\caption{Terms}
\label{fig:term}
\end{figure}
\subsection{Rules}
Each instruction is defined by a rule on the system model. The rule semantic is defined by several kinds of transitions:
\begin{enumerate}
\item \ExprTrans\ single-step evaluation of an expression in a system state,
\item \StateTrans\ internal transitions of a state, which is non-deterministic  as shown for loop instructions, such as \LOOP\ and \LOOPLEFT,
\item \SystemTrans\ symbolic system transitions.
\end{enumerate}

@@

\begin{mathpar}
\inferrule[]
  { \NEG\ \PREDICATE
  }{
  \{[\INSTRUCTION, \STACK, \PREDICATE]\} \cup \SYSTEM \SystemTrans \SYSTEM}
\end{mathpar}
Michelson instructions are categorized based on their functions. These include such as control structures, stack manipulation, arithmetic operations, boolean operations, cryptographic operations, and operations on data structures. In our model, we classified them as one-step, multi-step, blockchain, cryptographic, branch, and loop instructions. This categorization reflects how these instructions are handled within our specific model.
\subsubsection{One-step Instructions}
One-step instructions have a localized effect, modifying the system state directly without branching a new state, and they end the execution after a single operation. The ABS rule exemplifies such an instruction, operating on the first element of the stack. This element must be of type int, and the instruction returns the positive value of its current value, treated as a natural number.
%ABS
\begin{mathpar}
\inferrule[ABS]
  {
  }
  {
    [(\ABS; \INSTRUCTION), (\StackOne, \TINT) \STACKCONCAT \STACK,
    \PREDICATE]
    \StateTrans \\
    [\INSTRUCTION, (\X, \TNAT) \STACKCONCAT \STACK,
    \PREDICATE\ \wedge\ (\StackOne\ \GE\ \ZERO\ \Rightarrow\ \X\ =
    \StackOne) \wedge\ (\StackOne\ \LT\ \ZERO\ \Rightarrow\ \X\ = \MINUS\ \StackOne)]
 }
\end{mathpar}
Another example is the \COMPARE\ instruction for natural numbers, which compares the first two elements of the stack, both having a natural number type. It returns a result of either -1, 0, or 1 based on the comparison of their values.
% COMPARE
\begin{mathpar}
\inferrule[COMPARE]
  {
  }
  {
    [(\COMPARE ; \INSTRUCTION), (\StackOne, \TNAT) \STACKCONCAT (\StackTwo, \TNAT)
    \STACKCONCAT \STACK, \PREDICATE ]
    \SystemTrans \\
    [\INSTRUCTION, (\X, \TINT) \STACKCONCAT \STACK, \PREDICATE
    \wedge\ (\StackOne\ \GT\ \StackTwo\ \Leftrightarrow\ \X\ \EQ\ \ONE)
    \wedge\ (\StackOne\ \EQ\ \StackTwo\ \Leftrightarrow\ \X\ \EQ\ \ZERO) 
    \wedge\ (\StackOne\ \LT\ \StackTwo\ \Leftrightarrow\ \X\ \EQ\ \MINUS \ONE)]
    }
\end{mathpar}
\subsubsection{Multi-step Instructions}
Multi-step instructions involve the execution of a sub-sequence of instructions before returning to the main execution. The rule below models the \EXEC\ instruction, which executes a function specified as a sequence of instructions denoted by \INSTRUCTIONONE\ in the rule. This instruction applies the code of the function to the first element of the stack and then places the result back into the main execution.
\begin{mathpar}
  \inferrule[\EXEC]
  {
    [\INSTRUCTIONONE, (\StackOne, \TYF) \STACKCONCAT \EMPTYSTACK, 
    \Q]
    \StateTrans^*
    [ \EMPTYSTACK, (\StackOne', \TYS) \STACKCONCAT \Q']
  }{
     [(\EXEC; \INSTRUCTION),   (\{\INSTRUCTIONONE\}, \TYF\ \rightarrow\ \TYS) \STACKCONCAT
    (\StackOne, \TYF) \STACKCONCAT \STACK, \PREDICATE \wedge \Q] 
    \StateTrans \
    [ \INSTRUCTION, (\StackOne', \TYS) \STACKCONCAT \STACK,
    \PREDICATE \wedge \Q']}
\end{mathpar}

The \DIP\ \N\ instruction delves into the first \N\ elements of a stack, applies the sequence of instructions denoted as \INSTRUCTIONONE\ to the remaining part of the stack, modifies a part of the predicate and subsequently pushes the result back into the main execution.
%DIP n
\begin{mathpar}
\inferrule[\DIP\ \N]
  { 
     \FLEN(\A) \EQ\ \N \\ [\INSTRUCTIONONE,  \B, \Q]
    \StateTrans^*
    [\EMPTYSTACK,  \B_1, \Q']
  }
  {[(\DIP\ \N\ \INSTRUCTIONONE; \INSTRUCTION), \A\ \At\ \B, \PREDICATE \wedge \Q] \StateTrans 
[(\INSTRUCTION), \A\ \At\ \B_1, \PREDICATE \wedge \Q']}
\end{mathpar}
\subsubsection{Blockchain Instructions}
 Blockchain instructions make implicit use of data from an execution environment, derived from the current state of the blockchain (the context), from the transaction that triggered the contract, and from the block containing this transaction. We model these data as symbolic values, which satisfis some constrants. For example, the \AMOUNT\ instruction adds to the stack the amount of tokens that sent together with the transcation which calls this smart contract. Here, the amount is modeled as a symbolic value donated as \CAMOUNT, which has \TMUTEZ\ type, which has a nature number value.

\begin{mathpar}
\inferrule[AMOUNT]
  {
  }
  {[(\AMOUNT; \INSTRUCTION), \STACK, \PREDICATE] \StateTrans \
[\INSTRUCTION, (\CAMOUNT, \TMUTEZ) \STACKCONCAT \STACK, \PREDICATE]}
\end{mathpar}

\subsubsection{Cryptographic Instructions}
Cryptographic instructions in Michelson utilize built-in functions that manage cryptographic features. These functions are symbolically modeled in our framework, specifying their inputs and outputs. For instance, the instruction \SHA\ computes the cryptographic hash of the top of the stack using the SHA-256 cryptographic hash function. The function SHA-256 is symbolically represented as the function  \FSHA\  that takes an input of type \TBYTE\ and returns another byte value, constrained to a length of 256.
%HASH-KEY
\begin{mathpar}
\inferrule[\SHA]
  {
  }
  {[(\SHA; \INSTRUCTION), (\StackOne, \TBYTE) \STACKCONCAT\STACK, \PREDICATE] \StateTrans \
[\INSTRUCTION, (\X, \TBYTE) \STACKCONCAT\STACK, \PREDICATE \Wedge\ (\X\ \EQ\ \FSHA(\StackOne))]}
\end{mathpar}
\subsubsection{Branch Instructions}
The Michelson language incorporates conditional branching, enabling the execution of specific code segments based on certain criteria. One of the instructions facilitating this is the \IF\ \{\} \{\} construct. For example, the \IF\ \{ \INSTRUCTIONONE\ \} \{ \INSTRUCTIONTWO\  \} instruction allows the creation of branches in execution. It takes two sequences as arguments. Expecting a boolean value at the top of the stack, it consumes this top element. If the boolean is True, it executes the first provided sequence; otherwise, it executes the second sequence.
%IF
\begin{mathpar}
  \inferrule[IF-true]
  {  
  }{
    [(\IF\ \INSTRUCTIONONE\  \INSTRUCTIONTWO; \INSTRUCTION),
    (\StackOne, \TBOOL) \STACKCONCAT\STACK, \PREDICATE]
    \StateTrans\
    [\INSTRUCTIONONE, \STACK, \PREDICATE\ \Wedge\ \StackOne]
  }

@@

Another branch instruction is the IF-LEFT instruction, which is used for conditional execution based on the left or right branch of a tagged union (sum) type. 
%IF-LEFT-LEFT
\begin{mathpar}
  \inferrule[IF-LEFT-left]
  {  
  }{
    [(\IFLEFT\ \INSTRUCTIONONE\ \INSTRUCTIONTWO; \INSTRUCTION),
    (\StackOne, \TOR\ \TYF\ \TYS) \STACKCONCAT \STACK, \PREDICATE]
    \StateTrans \
    [\INSTRUCTIONONE, (\X, \TYF) \STACKCONCAT\STACK,
    \PREDICATE \wedge (\StackOne\ \EQ\ \LEFT\ \X)]
  }
\end{mathpar}

%IF-LEFT-RIGHT
\begin{mathpar}
  \inferrule[IF-LEFT-right]
  {  
  }{
    [(\IFLEFT\ \INSTRUCTIONONE\  \INSTRUCTIONTWO; \INSTRUCTION),
    (\StackOne, \TOR\ \TYF\ \TYS) \STACKCONCAT \STACK, \PREDICATE]
    \StateTrans \
    [\INSTRUCTIONTWO, (\X, \TYS) \STACKCONCAT\STACK, \PREDICATE \wedge (\StackOne\ \EQ\ \RIGHT\ \X))]
  }
\end{mathpar}
These above rules collectively define the behavior of the \IFLEFT instruction based on the left or right branch of a tagged union type. This rule applies when the top element of the stack is a tagged union of type \TOR\ \TYF\ \TYS (either of type \TYF\ or \TYS), and the \IFLEFT instruction is followed by instructions \INSTRUCTIONONE\ and \INSTRUCTIONTWO. The first rule apply with the condition that the top element of the stack is of the form Left \X\ (indicating the left branch of the union), it transitions to execute the instructions \INSTRUCTIONONE\ with the stack modified accordingly and the predicate is updated to include the condition that the top element of the stack is equal to Left \X.  Similerly, the second rule apply with the condition that  the top element is of the form Right \X\ (indicating the right branch of the union), it transitions to execute the instructions \INSTRUCTIONTWO\ with the stack modified accordingly and the predicate is updated to include the condition that the top element of the stack is equal to Right \X.

\subsubsection{Loop Instructions}
In Michelson, loops are fundamental constructs that repeatedly execute a set of instructions until a specific condition is met. There are various loop instructions, including \ITER, \MAP, \CONCAT, \SIZE, \LOOP, and \LOOPLEFT, enable the manipulation and traversal of different data structures like lists, sets, maps, and the stack itself.

Michelson, similar to other programming languages, classifies loop instructions into two main types:
\begin{itemize}
\item \KFOR\ loops: These are sequence-based loops, exemplified by \ITER, \MAP, \CONCAT\ and \SIZE.
\item \KWHILE\ loops: These are predicates-based loops, illustrated by \LOOP and \LOOPLEFT.
\end{itemize}
To illustrate the structures of  loops, we can draw a parallel with OCaml syntax:
\begin{verbatim}
while <loop_condition> do
   <loop_body>
done
\end{verbatim}
This \KWHILE\  loop structure involves a loop condition that is evaluated at the beginning of each iteration. If the condition holds true, the loop body is executed, and the process repeats until the condition becomes false.  In Michelson, this pattern is evident in instructions like LOOP and LOOP-LEFT, which rely on predicate-based conditions for their iteration.
\begin{verbatim}
LOOP {<loop-body>}
\end{verbatim}
Additionally, Michelson features a \KFOR\ loop, where the index runs from the initial value to the final value:
\begin{verbatim}
for index = <initial_value> to <final_value> do
  <loop_body>
done
\end{verbatim}
This is represented by the ITER instruction:
\begin{verbatim}
ITER {<loop-body>}
\end{verbatim}
